Numerical modelling of the Cordilleran ice sheet

### 2014-08-25 kappa.tex
Add numerical implementation paragraph.

A final integration over the entire ice column yields
\citep[Eqs.~7.7--7.8]{Weis.etal.1999}
\begin{subequations}
\begin{gather}
  \label{eqn:vssa}
  \begin{split}
    \pdv{x} \left[2\bar{\nu}h
                  \left(2\pdv{v_x}{x} + \pdv{v_y}{y}\right) \right]
        &+ \pdv{y} \left[\bar{\nu}h
                         \left(\pdv{v_x}{y} + \pdv{v_y}{x}\right) \right] \\
        &+ \tau_{bx} = \rho gh \pdv{s}{x} \,,
  \end{split}\\
  \begin{split}
    \pdv{x} \left[\bar{\nu}H
                  \left(\pdv{v_x}{y} + \pdv{v_y}{x}\right) \right]
        &+ \pdv{y} \left[2\bar{\nu}h
                         \left(\pdv{v_x}{x} + 2\pdv{v_y}{y}\right) \right] \\
        &+ \tau_{by} = \rho gh \pdv{s}{y} \,,
  \end{split}
\end{gather}
\end{subequations}
where $\tau_{bx}$ and $\tau_{by}$ correspond to $x$ and $y$-components of the
basal traction, and will be determined by a sliding law.

@@

% ----------------------------------------------------------------------
%\subsection{Numerical implementation}
% ----------------------------------------------------------------------

%\todo{Add here a short section about discretization of the SIA and SSA, and
%      their numerical implementation. The main subtlety, I think, lies in the
%      fact that some computation need to use a ``staggered'' grid in order to
%      ensure that we add SIA and SSA values corresponding to the same point in
%      space. I need to check that. This section is also the place where to
%      introduce geographic transformation parameters that I used throughout
%      the thesis -- except in \cref{fig:map-cordillera}!}

%\begin{table}
%  \caption{Properties of the Canadian Atlas Lambert conformal conic projection
%           (EPSG code~3978) and the GRS~1980 ellipsoid, and corresponding
%           PROJ.4 parameter names, as used in all model runs, ice volume and
%           area calculations, and most figures.}
%  \label{tab:params}
%  \begin{tabular*}{80mm}{@{\extracolsep{\fill}}lcc}
%    \hline
%    Parameter               & Name              & Value \\
%    \hline
%    Projection name         & \texttt{+proj}    & \texttt{lcc} \\
%    Central meridian        & \texttt{+lon\_0}  & 95\degree W \\
%    Latitude of origin      & \texttt{+lat\_0}  & 49\degree N \\
%    First secant parallel   & \texttt{+lat\_1}  & 49\degree N \\
%    Second secant parallel  & \texttt{+lat\_2}  & 77\degree N \\
%    Semimajor radius        & \texttt{+a}       & 6378137\,m \\
%    Reciprocal flattening   & \texttt{+rf}      & 298.257222101 \\
%    \hline
%  \end{tabular*}
%\end{table}

% ======================================================================
%\section{Software implementation}
% ======================================================================

% ----------------------------------------------------------------------
%\subsection{Computational workflow}
% ----------------------------------------------------------------------
%\todo{Figure. computational workflow}

% ----------------------------------------------------------------------
%\subsection{Contributions to PISM}
% ----------------------------------------------------------------------
%\todo{Table. PISM ``issues'' identified}

% ----------------------------------------------------------------------
%\subsection{Peripheral tools}
% ----------------------------------------------------------------------
%A Python PDD model\\
%Pre-processing scripts for GRASS GIS\\
%Plotting tools for PISM\\

>>>>

A final integration over the entire ice column yields
\citep[Eqs.~7.7--7.8]{Weis.etal.1999}
\begin{subequations}
\label{eqn:vssa}
\begin{gather}
  \begin{split}
    \pdv{x} \left[2\bar{\nu}h
                  \left(2\pdv{v_x}{x} + \pdv{v_y}{y}\right) \right]
        &+ \pdv{y} \left[\bar{\nu}h
                         \left(\pdv{v_x}{y} + \pdv{v_y}{x}\right) \right] \\
        &+ \tau_{bx} = \rho gh \pdv{s}{x} \,,
  \end{split}\\
  \begin{split}
    \pdv{x} \left[\bar{\nu}H
                  \left(\pdv{v_x}{y} + \pdv{v_y}{x}\right) \right]
        &+ \pdv{y} \left[2\bar{\nu}h
                         \left(\pdv{v_x}{x} + 2\pdv{v_y}{y}\right) \right] \\
        &+ \tau_{by} = \rho gh \pdv{s}{y} \,,
  \end{split}
\end{gather}
\end{subequations}
where $\tau_{bx}$ and $\tau_{by}$ correspond to $x$ and $y$-components of the
basal traction, and will be determined by a sliding law.

@@

% ----------------------------------------------------------------------
\subsection{Numerical implementation}
% ----------------------------------------------------------------------

In PISM, the SIA and the SSA are resolved numerically using centred, explicit
finite difference methods. Ice velocities, $\vsia$ and $\vssa$, are computed on
a staggered grid, i.e., a grid offset by half a grid cell in both directions of
the $x$- and $y$-axes \citep{Bueler.Brown.2009, Winkelmann.etal.2011}. The
computation of SSA velocities involves the numerical resolution of
\cref{eqn:vssa}. After discretization, this can be thought of as the inversion
of a large but sparse linear system of the form $\tens{A}\vssa=\vect{b}$. This
sparse linear problem is resolved in parallel by numerical iteration methods
implemented in PETSc \citep{Balay.etal.2014}. The simulations presented here
use the generalized minimal residual method \citep[GMRES,][]{Saad.2003}. The
vertically-averaged mass continuity equation is then treated as a
diffusion-advection equation, where SIA velocities are diffusive and SSA
velocities advective \citep{Bueler.Brown.2009, Winkelmann.etal.2011}. The
horizontal grid uses the Canadian Atlas Lambert conformal conic projection
(\cref{tab:proj}). Time-stepping is adaptive.

\begin{table}
  \caption{Properties of the Canadian Atlas Lambert conformal conic projection
           (EPSG code~3978) and the GRS~1980 ellipsoid, and corresponding
           PROJ.4 parameter names, as used in all model runs, ice volume and
           area calculations, and most figures.}
  \label{tab:proj}
  \begin{tabular*}{80mm}{@{\extracolsep{\fill}}lcc}
    \tophline
    Parameter               & Name              & Value \\
    \middlehline
    Projection name         & \texttt{+proj}    & \texttt{lcc} \\
    Central meridian        & \texttt{+lon\_0}  & 95\degree W \\
    Latitude of origin      & \texttt{+lat\_0}  & 49\degree N \\
    First secant parallel   & \texttt{+lat\_1}  & 49\degree N \\
    Second secant parallel  & \texttt{+lat\_2}  & 77\degree N \\
    Semimajor radius        & \texttt{+a}       & 6378137\,m \\
    Reciprocal flattening   & \texttt{+rf}      & 298.257222101 \\
    \bottomhline
  \end{tabular*}
\end{table}
